---
title: "Homework 2"
author: "Fengdi Zhang"
date: "2022-10-04"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
```

## Problem 1

Load the NYC Transit Data
```{r, message=FALSE}
nyc_transit = 
      read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
      janitor::clean_names() %>%
      select(line, station_name, station_latitude, station_longitude, 
             starts_with("route"), entry, vending, entrance_type, ada) %>% 
      mutate(entry = as.logical(recode(entry, YES = "TRUE", NO = "FALSE")))

```

The csv files has loaded and the variable names are cleaned. Only desired variabels were retained, including `r ls(nyc_transit)`. The dimension of the resulting dataset is `r nrow(nyc_transit)` rows and `r ncol(nyc_transit)` columns. The data is not tidy since the some variables can be combines, such as the routes. 

* There are `r nrow(distinct(nyc_transit, line, station_name))` distinct stations.
* There are `r sum(nyc_transit$ada)` stations that are ADA compliant.
* The proportion of station entrances/exits without vending allow entrance is `r (length(nyc_transit$entry[(nyc_transit$entry ==TRUE) & (nyc_transit$vending == "NO")]))/length(nyc_transit$vending[nyc_transit$vending == "NO"])`. 

Reformat data so that route number and route name are distinct variables.
```{r}
nyc_transit = nyc_transit %>% 
  mutate(
         route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) %>% 
  pivot_longer(
         route1:route11,
         names_to = "route_number", 
         names_prefix = "route",
         values_to = "route_name")
```

There are `r nrow(distinct(filter(nyc_transit, route_name =="A"), station_name))` distinct stations serve the A train. Among the stations that serve the A train, `r nrow(distinct(filter(nyc_transit, route_name =="A", ada == TRUE), station_name))` are ADA compliant. 



## 


```{r echo=FALSE}

```

